Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{{DISTRIBUTION STATEMENT A. Approved for public release.} }

### 2021-05-13 paper/agree.tex

The goal of this section is to illustrate in more detail the process a system designer follows to add cyber requirements to the AGREE specifications and then automatically transform the system to insert the high-assurance components. The AGREE specifications generated by the transforms for the high-assurance components are explained. The section ends with a concise formal statement of the meaning of an AGREE specification for a high-assurance component. This meaning is what must be preserved by synthesis.

@@

The AGREE specification for the SW component in the example with the added cyber requirements is given in \figref{fig:sw}. Without going into too much detail, the language is a first order predicate calculus. The semantics are synchronous data-flow where the inputs, outputs, and expressions are characterized by data streams that comply with the input assumptions. The semantics are such that the contracts are evaluated in dependency order with inputs being propagated to outputs through all the contracts until they stabilize; as such, the contracts, and thereby the top-level model, must be acyclic. Once the contracts have stabilized, then the model takes a synchronous step to the next input data in the stream. The semantics do not model computation or communication delay. The output of one contract is seen at the input of any downstream contract in the same step of the input data stream.

The AGREE model checker attempts to prove several properties of the top-level model being verified. The first is that the output guarantees of each component implementing the system are strong enough to guarantee the input assumptions of any downstream component and strong enough to satisfy the guarantees of the top-level component's output that is being verified (i.e., the system composition meets input assumptions at each input and the guarantees on the system output). These properties are reported in the expanded lists in \figref{fig:example}(b) and \figref{fig:hardened}(b).  The next set of properties prove that the contract specifications for each component are self-consistent (i.e, a contract does not contradict itself). These are the unexpanded results to the same two figures.

Returning back to the contract in \figref{fig:sw}, it uses \texttt{eq} statements to define variables local to the contract specification. For example, the \texttt{req} variable is equivalent to the \texttt{event(AutomationRequest)} expression. In the AGREE semantics, there is an implicit \emph{event} input (or output) associated with every named event port in a component. The semantics used here do not not buffer these events so the implicit input (or output) is only a boolean value. An \texttt{event} expression refers to that implicit input (or output) and is true if there is data on the named port. The system contract here states assumptions on well-formed input, and it then states guarantees on properties about the output.

@@

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
      \includegraphics[scale=0.3]{dialogue.png}
    \end{tabular}
  \end{center}
  \caption{Dialogue for automatically transforming the model with a filter.}
  \label{fig:dialogue}
\end{figure}

@@

As noted previously, the original system fails to guarantee the cyber requirement. BriefCASE provides two transformations to address the failing requirements: inserting a filter and inserting a monitor. The component is added by selecting the connection in the model where the high-assurance component is to be added, and then choosing the appropriate transformation. The dialogue or a filter transformation is shown in \figref{fig:dialogue}. The system designer can state the policy directly in the dialogue, or leave it  blank. In this example, the policy is specified as \texttt{WELL\_FORMED\_AUTOMATION\_RESPONSE(Input)}. Additionally the filter can be attached to any requirement the vulnerability analysis. This association is useful the the assurance case from the Resolute tool [CITE DESTION 2021]. The \emph{OK} button creates the component and inserts it accordingly into the implementation

The AGREE contract specification generated by the transform is shown in \figref{fig:assurance}(a). The guarantee is stylized for synthesis and completely defines the meaning of the output under all possible input. As similar dialogue exist for inserting monitors that allows the system designer to add and remove inputs as needed. The resulting AGREE specification for the monitor in this example is shown in \figref{fig:assurance}(b). The \texttt{is\_latched} value makes the alert persistent meaning that once raised always raised and is one of the several options available in the dialogue. The definition for \texttt{policy} is taken by the system designer from the contract in \figref{fig:sw}. As before the guarantees for the outputs are autogenerated by the tool and completely define each output under every possible input.

>>>>

The goal of this section is to illustrate in more detail the process a system designer follows to add cyber requirements to the AGREE specifications and then automatically transform the system to insert the high-assurance components. The AGREE specifications generated by the transforms for the high-assurance components are explained. The section ends with a concise formal statement of the meaning of an AGREE specification for a high-assurance component. This meaning is what must be preserved by the synthesis.

@@

The AGREE specification for the SW component in the example with the added cyber requirements is given in \figref{fig:sw}. Without going into too much detail, the language is a first order predicate calculus. The semantics are synchronous data-flow where the inputs, outputs, and expressions are characterized by data streams that comply with the input assumptions. The semantics are such that the contracts are evaluated in dependency order with inputs being propagated to outputs through all the contracts until they stabilize; as such, the contracts, and thereby the top-level model, must be acyclic. Once the contracts have stabilized, the model takes a synchronous step to the next input data in the stream. The semantics do not model computation or communication delay. The output of one contract is seen at the input of any downstream contract in the same step of the input data stream.

The AGREE model checker attempts to prove several properties of the top-level model being verified. The first is that the output guarantees of each component implementing the system are strong enough to validate the input assumptions of any downstream component and strong enough to satisfy the guarantees of the output of the top-level component being verified (i.e., the system composition meets input assumptions at each input and the guarantees on the system output). These properties are reported in the expanded lists in \figref{fig:example}(b) and \figref{fig:hardened}(b).  The next set of properties prove that the contract specifications for each component are self-consistent (i.e, a contract does not contradict itself). These are the unexpanded results at the bottom of the figures.

Returning back to the contract in \figref{fig:sw}, it uses \texttt{eq} statements to define variables local to the contract specification. For example, the \texttt{req} variable is equivalent to the \texttt{event(AutomationRequest)} expression. In the AGREE semantics, there is an implicit \emph{event} input (or output) associated with every named event port in a component. The semantics used here do not buffer these events so the implicit input (or output) is only a boolean value. An \texttt{event} expression refers to that implicit input (or output) and is true when data is placed on the named port. The system contract here states assumptions on well-formed input, followed by guarantees on properties about the output.

@@

\begin{figure}
  \begin{center}
    \begin{tabular}{c}
      \includegraphics[scale=0.3]{dialogue.png}
    \end{tabular}
  \end{center}
  \caption{Wizard for automatically transforming the model with a filter.}
  \label{fig:dialogue}
\end{figure}

@@

As noted previously, the original system fails to guarantee the cyber requirement. BriefCASE provides two transformations to address the failing requirements: inserting a filter and inserting a monitor. The component is added by selecting the connection in the model where the high-assurance component is to be added, and then choosing the appropriate transformation. The system designer can provide transform configuration parameters in a wizard, as shown in \figref{fig:dialogue}. The policy of the high assurance component can be stated directly in the wizard, or leave it  blank. In this example, the policy is specified as \texttt{WELL\_FORMED\_AUTOMATION\_RESPONSE(Input)}. 
Additionally, because a transformation is ultimately driven by a cyber requirement, BriefCASE updates an embedded Resolute assurance case~\cite{resolute-destion}.  Resolute keeps track of the evidential artifacts necessary for supporting the requirement, and can be run anytime throughout the design phase to determine whether those artifacts are valid.
%Additionally, the filter can be attached to any requirement the vulnerability analysis. This association is useful the the assurance case from the Resolute tool [CITE DESTION 2021]. The \emph{OK} button creates the component and inserts it accordingly into the implementation

The AGREE contract specification generated by the transform is shown in \figref{fig:assurance}(a). The guarantee is stylized for synthesis and completely defines the meaning of the output under all possible input. 
%A similar dialogue exist for inserting monitors that allows the system designer to add and remove inputs as needed. 
The resulting AGREE specification for the monitor in this example is shown in \figref{fig:assurance}(b). The \texttt{is\_latched} value makes the alert persistent meaning that once raised always raised and is one of the several options available in the dialogue. The definition for \texttt{policy} is taken by the system designer from the contract in \figref{fig:sw}. As before the guarantees for the outputs are autogenerated by the tool and completely define each output under every possible input.

### 2021-05-13 paper/intro.tex

The two automatic transforms discussed in this paper are (1) the insertion of a filter to prevent malformed data from a malicious actor from being propagated to downstream components, and (2) the insertion of a monitor to detect and alert unexpected behaviors arising from untrusted components. These transformations not only change the architecture of the model by adding in new components, but they generate a complete formal specification for the behavior of the inserted high-assurance components in the AGREE language. Those specifications are sufficient for model checking to prove that with the high-assurance components, the hardened system meets its cyber-resiliency requirements.

The other novel aspect of the approach is the synthesis of the AGREE specifications for the high-assurance components to CakeML which is a verified implementation of ML \cite{cakeml}. This paper describes in detail the synthesis path from specifications to CakeML code, providing a formal framework in which to argue correctness. CakeML then provides a verified compilation path to several different target binaries proving that the meaning of the CakeML is exactly preserved in the final binaries. Assuming the deployed cyber-hardened system is scheduled as intended by the AADL model, and the HAMR generated communication fabric delivers messages between components as expected, then the AGREE model checking results hold for the deployed system, \ie, it detects and prevents the indicated cyber-vulnerabilities over all possible finite inputs. Future work is lifting this result to infinite input traces as these systems are inherently reactive and intended to run forever.

The approach is motivated, and illustrated, in a simple example. A full case study applying these transformations to a UAV system that uses the Air Force Research Laboratory's OpenUxAS services for route planning is briefly reported. Here the transforms add filters to guard against malformed input and monitors to guard against ground station spoofing and malicious flight plans from OpenUxAS. The case study system is significantly more complex than the example and shows the viability of the modeling approach to full-scale industrial design.

>>>>

The two automatic transforms discussed in this paper are (1) the insertion of a filter to prevent malformed data from a malicious actor from being propagated to downstream components, and (2) the insertion of a monitor to detect and alert unexpected behaviors arising from untrusted components. These transformations not only change the architecture of the model by adding in new components, but they generate a formal specification for the behavior of the inserted high-assurance components in the AGREE language. Those specifications are sufficient for model checking to prove that with the high-assurance components, the hardened system meets its cyber-resiliency requirements.

The other novel aspect of the approach is the synthesis of the AGREE specifications for the high-assurance components to CakeML, which is a verified implementation of ML \cite{cakeml}. This paper describes in detail the synthesis path from specifications to CakeML code, providing a formal framework in which to argue correctness. CakeML then provides a verified compilation path to several different target binaries proving that the meaning of the CakeML is exactly preserved in the final binaries. Assuming the deployed cyber-hardened system is scheduled as intended by the AADL model, and the HAMR generated communication fabric delivers messages between components as expected, then the AGREE model checking results hold for the deployed system, \ie, it detects and prevents the indicated cyber-vulnerabilities over all possible finite inputs. Future work is lifting this result to infinite input traces as these systems are inherently reactive and intended to run forever.

The approach is motivated, and illustrated, in a simple example. A case study applying these transformations to a UAV system that uses the Air Force Research Laboratory's OpenUxAS services for route planning is presented. Here the transforms add filters to guard against malformed input and monitors to guard against ground station spoofing and malicious flight plans from OpenUxAS. The case study system is significantly more complex than the example and shows the viability of the modeling approach to full-scale industrial design.
